---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# openalexR

<!-- badges: start -->
<!-- badges: end -->

The goal of openalexR is to create an R wrapper for the Open Alex API (https://openalex.org) which allows users to query and extract results from the Open Alex database of scholarly output. Open Alex succeeds the Microsoft Academic Graph in maintaining an extensive relational database. The API has 5 endpoints to query: (1) Works (journal & conference papers, books, data, etc); (2) Authors; (3) Venues; (4) Institutions; and (5) Concepts. More information about each of these endpoints can be found here: (https://docs.openalex.org). 

The API is completely free, has no rate limits, and is open source. If you prefer, you can also request a snapshot of all of the data to be copied into an amazon s3 bucket (https://docs.openalex.org/download-snapshot). 

openalexR has functions to request data from each of the five endpoints: get_works, get_authors, get_venues, get_institutions, and get_concepts. It also has some wrapper functions for specific needs (get_authors_papers, get_authors_coauthors). 

## Installation

You can install the development version of openalexR from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ekmaloney/openalexR")
```

## EWorks 

The get_works function takes three main parameters: 
(1) ID type, one of: doi, open_alex, mag (Microsoft Academic Graph), pmid (Pubmed Identifier), pmcid (Pubmed Central Identifier)
(2) Id(s), can either be 1 id or a list of ids. 
(3) Variable_unnest (optional) - a variable by which you would like to make the dataframe longer (can be: authors, )

```{r example}
library(openalexR)
library(tidyverse)

#find a paper in the openalex database: 

#Example: "Analysis of the glottal excitation of emotionally styled and stressed speech,"
#one of my mother's publications when she was an electrical engineer at Georgia Tech
#I will use the doi to retrieve it: 

paper_info <- find_work(id_type = "doi", 
                         id = "https://doi.org/10.1121/1.413664",
                         variable_unnest = "authors")

glimpse(paper_info)

```

From the paper info, you can also find all papers by an author: 

```{r cars}

mom_oa_id <- paper_info$authors$author_id[1]

all_papers <- get_authors_papers(id_type = "openalex",
                                 id = mom_oa_id)


glimpse(all_papers)
```

